---
title: "Expense Analysis"
author: "Le Thi Thanh Tam"
date: "8/31/2022"
output: pdf_document
---

## Objectives

* For this funny project, I just want to quickly analyze my own spending habit based on my personal bank account data.
Since I realize that I am not good at personal financial management, figure out how much I spend per month to help me manage my money and save my life (:.

* The data is stored in two years from Sep 2020 to August 2022 _ The time I've been to Korea.

## Data Pre-processing

```{r}
# library packages
library(readxl)
library(data.table)
library(ggplot2)
library(dplyr)

# Read data
Sys.setlocale("LC_ALL","C")
dt_in <- read_xls("E:/ThanhTam_DA/Project/BankTransaction.xls")
Sys.setlocale("LC_ALL", "Korean")
glimpse(dt_in)
colnames(dt_in)
# Let's say I just want to focus on my spendings amount so I will delete some unnecessary columns 
# ("Deposit Amount", "Transaction Time", "Transaction Place")
dt_dropped <- dt_in[,-(5:7)]
# Also drop values equal 0 in 'Withdrawal Amount' column (it means these transactions is for 'Deposit Amount')
dt <- dt_dropped[!(dt_dropped$`Withdrawal Amount` == 0),]

# Rename column
colnames(dt) = c("Date", "Category", "Name", "Amount")

# Next Format the Date type
# Add the Month and Year column to the data
dt$Date <- as.Date(as.character(dt$Date), format = "%Y%m%d")
dt$Month <- format(as.Date(dt$Date), "%m")
dt$Year <- format(as.Date(dt$Date), "%Y")

# Check missing values
sum(is.na(dt)) # The numbers of NA values is too small we can delete 4 NA values
dt[which(is.na(dt)),] 
dt <- na.omit(dt)
dt$Category = NULL
head(dt)
```

## Analysis

### 1. Fixed Expenses

* First I want to see the average fixed expenses I spent per month
* Fixed amount include house rent, phone bills, monthly transportation fee and donation

```{r}
# By identify repeated name by month
# Count values by column 'Name' and 'Amount'
dt <- data.table(dt)
is.data.table(dt)
dt[, freq := .N, by = .(Name, Amount)] # count by Name

# Assign values have frequency bigger than 1 is a fixed amount
dt[freq>1, type := "fixed"]
# Create new data frame with only those fixed values
dt_fixed <- dt[type=="fixed"]

# Let have a look at the  new data
summary(dt_fixed)

# Sort the data in increasing order by column freq
dt_fixed_sorted <- dt_fixed[order((dt_fixed$freq),decreasing = TRUE), ]
head(dt_fixed_sorted, n=20) # I wanna see what catogory I spent most frequently (20 times) --> it turns out the convinient store near my last company (: 

## It seems like including more others variable expenses
## So I need to classify and assign fixed expense only catogories including: house rent, phone bills, bank transportation fee and donation
#------------------------------

# Create a new data frame with adding the frequency column by 'Name' column
dt[, freq := .N, by = .(Name)] # Count by Name
data <- dt[freq>1]
data$type = NULL

##1. house rent
# Assign 'house rent' values to Category column
rent.key <- c("김사묵", "NGUYEN NGO", "허길행","한전(김명섭)","한전(김순묵)")
rent.key.pattern <- paste(rent_key, collapse = "|")
data[grepl(rent_key_pattern, Name), Category := "house rent"]

##2. phone bills
# Assign 'phone bills' values to Catogory column
phone.bill <- c("LGU+통신요금자동","LGU+LE THI","KT7906159606","KT7906159604", "KT7906159608","KT7906159601", "KT7906159603", "KT7906159607", "KT7906159609","통신판매(ARS)_나")
phone.bill.pattern <- paste(phone_bill, collapse = "|")
data[grepl(phone_bill_pattern, Name), Category := "phone bill"]

#3. Transportation fee
# Assign monhly 'transpotation' fee to Category column
trans.key <- c("하나카드_체크","레티따인떰(하나카드)","하나카드")
trans.key.pattern <- paste(trans_key, collapse = "|")
data[grepl(trans_key_pattern, Name), Category := "transportation"]

#4. donation
# Asign 'donation' value to Category column
data$Category[data$Name == "후원금구(1)_나이"] = "donation"


# Now Create a data frame for fixed expenses above and compute how much spendings I did by month per year
data_fixed_exp <- data[(data$Category == "house rent"| data$Category == "phone bill"| 
                          data$Category == "transportation"|data$Category =="donation"),]

# The total fixed amount I spent per month
FixedExp_by_month <- data_fixed_exp[, .(FixedExpense = sum(Amount)), by = .(Month=paste(Month, "/", Year))]

# Vizualization
FixedExp_by_month[, Month := factor(Month, levels = Month)]

ggplot(data = FixedExp_by_month, aes(x=Month, y = FixedExpense)) +
  geom_bar(stat = "identity", fill = "#219ebc") +
  geom_hline(aes(yintercept = mean(FixedExpense), color="red")) +
  theme(axis.text.x = element_text(size = 5,angle = 45, vjust = 0.5, hjust=1))

# It looks like my fixed expenses varied much --> i want to see more in details
#-----------------------------
# Classify by each category to see spendings in each one
FixedExp_by_category <- data_fixed_exp[, .(FixedExpense = sum(Amount)), 
                                       by = .(Month=paste(Month, "/", Year), Category)]

FixedExp_by_category[, Month := factor(Month, levels = unique(Month))]

ggplot(data = FixedExp_by_category, aes(fill = Category,y = FixedExpense, x =Month)) +
  geom_bar(position="stack", stat="identity") +
  ggtitle("Fixed Expenses of each Category by Month") +
  geom_hline(aes(yintercept = mean(FixedExpense), color="red", show_guide = FALSE)) +
  theme(axis.text.x = element_text(size = 5,angle = 45, vjust = 0.5, hjust=1))

# As expected, 'house rent' costs the highest amount and it did not remain contant since I also paid for deposit house rent and advance payment
# But in general the average fixed amount that i need to pay some how under 250,000 KRW . It is a good sign (:
# Anyway I may consider my T-money Card and also paid phone fees have increased dramatically last month to decrease these costs.

ex <- data[order((dt_fixed$Amount),decreasing = TRUE),]
a <- data.frame(table(ex$Name))
a[order((a$Freq),decreasing=TRUE),]
```

### 2. Variable Expenses

* Variable Expenses are known as expenses which change from month to month (dining out, medical expenses, groceries or anything buy from a store). By understand my spending behaviour, i could lower those.

### 2.1. Food Expense

* I want to see How much I spent for eat out per month

```{r}
food_key <- c("큰맘할매순대국우", "요기요(제휴카드)", "쥬봉쌀국수", "텐더(TENDER)",
              "솔뫼해장국","윤가네옛날통닭&","KFC가산디지털점", "빵이가", "AKIMOTO MA",
              "가마솥서울깍두기", "손큰할매순대국", "연이자카야")
food_key_pattern <- paste(food_key, collapse = "|")
data[grepl(food_key_pattern, Name), Category := "food"]

# Compute
FoodExp_by_month <- data[data$Category == "food", .(FoodExpense = sum(Amount)), by = .(Month=paste(Month, "/", Year))]

# Vizualization
FoodExp_by_month[, Month := factor(Month, levels = Month)]

ggplot(data = FoodExp_by_month, aes(x=Month, y = FoodExpense)) +
  geom_bar(stat = "identity", fill = "#6495ED") +
  ggtitle("Food Expense") +
  ylim(0,100000) + 
  geom_hline(aes(yintercept = mean(FoodExpense))) + 
  theme(axis.text.x = element_text(size = 5,angle = 45, vjust = 0.5, hjust=1))

# I tended to decrease my habit to eat outside. Good !


```

### 2.2 Cafe Expense

```{r}
cafe_key <- c( "투썸플레이스서린", "팔공티우송대학교", "공차우송대점"
             ,"투썸플레이스서울", "스타벅스코리아", "팔공티은행점",
              "공차", "스타벅스커피코리", "공차대전복합터미", "만랩커피신설동역")
cafe_key_pattern <- paste(cafe_key, collapse = "|")
data[grepl(cafe_key_pattern, Name), Category := "cafe"]
data$Category[data$Name == "카페1800(cafe180" ] = "cafe"
data$Category[data$Name == "redLcoffee(레드" ] = "cafe"

# Compute

CafeExp_by_month <- data[data$Category == "cafe", .(CafeExpense = sum(Amount)), by = .(Month=paste(Month, "-", Year))]

# Vizualization
CafeExp_by_month[, Month := factor(Month, levels = Month)]

ggplot(data = CafeExp_by_month, aes(x=Month, y = CafeExpense)) +
  geom_bar(stat = "identity", fill = "#fb8500") +
  ylim(0,200000) + 
  ggtitle("Cafe Expense") +
  geom_hline(aes(yintercept = mean(CafeExpense))) +
  theme(axis.text.x = element_text(size = 5,angle = 45, vjust = 0.5, hjust=1))


```


